Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-08-03 submission/manuscript.Rmd
include Ada suggestions results 1st 2 paragraphs

**Decreased SCFA concentrations are not associated with tumors.** We used high-performance liquid chromatography (HPLC) to measure acetate, butyrate, and propionate concentrations of frozen fecal samples from `r length(metai$sample)` individuals at a cross-sectional point in time. There was no difference between individuals without colon tumors (n=`r length(filter(metai, dx == "normal") %>% pull(sample))`) and patients with either an adenoma (n=`r length(filter(metai, dx == "adenoma") %>% pull(sample))`) or carcinoma (n=`r length(filter(metai, dx == "cancer") %>% pull(sample))`) for any of the SCFAs measured after multiple comparison correction (P-value `r ifelse(min(pull(hplc_groups_scfa_kruskal, bh)) > 0.15, invisible(">"), invisible("<"))` 0.15) [Figure 1A - 1C]. We next measured the concentration of SCFAs in `r length(metaf$initial)` patients with an adenoma (n=`r length(filter(metaf, dx == "adenoma") %>% pull(initial))`) or carcinoma (n=`r length(filter(metaf, dx == "cancer") %>% pull(initial))`) in which we had pre- and post-treatment fecal samples. Although there was a general trend for increasing acetate, butyrate, and propionate concentrations after treatment for tumors, there was no significant difference pre- and post-treatment for patients with adenomas (P-value `r ifelse(min(filter(hplc_treat_scfa_kruskal, disease == "adenoma") %>% pull(bh)) > 0.20, invisible(">"), invisible("<"))` 0.20) or carcinomas (P-value `r ifelse(min(filter(hplc_treat_scfa_kruskal, disease == "carcinoma") %>% pull(bh)) > 0.80, invisible(">"), invisible("<"))` 0.80) [Figure 1D - 1F].

**Changes in genes for enzymes involved in SCFA synthesis are not associated with tumors.** Using a list of specific genes that are important for the synthesis of SCFAs [Table S1], we looked for differences in gene abundance between individuals without colon tumors and patients with adenomas or carcinomas. First, using imputed gene relative abundance based on 16S rRNA gene sequencing we found no difference in any of the genes involved with acetate, butyrate, or propionate synthesis (P-value `r ifelse(min(pull(picrust_select, bh)) > 0.90, invisible(">"), invisible("<"))` 0.90) [Table S2]. This similarity between groups is highlighted by visualizing genes important in butyrate synthesis [Figure 2A]. Using a paired Wilcoxon rank-sum test, there also was no difference in imputed gene relative abundance between pre- and post-treatment samples for any genes involved with SCFA synthesis (P-value `r ifelse(min(pull(picrust_select_treat, bh)) > 0.70, invisible(">"), invisible("<"))` 0.70) [Table S3]. Next, we took a subset of these `r length(metai$sample)` fecal samples (n=`r length(geof_samples$X1)`) and used metagenomic sequencing to confirm these results. Like the imputed gene results, metagenomic analysis found that there was no difference in any of the genes involved in SCFA synthesis between individuals without colon tumors (n=`r length(filter(geof_samples, X30 == "Healthy") %>% pull(X2))`) and patients with adenoma (n=`r length(filter(geof_samples, X30 == "Adenoma") %>% pull(X2))`) or carcinoma (n=`r length(filter(geof_samples, X30 == "Cancer") %>% pull(X2))`) (P-value `r ifelse(min(pull(opf_select, bh)) > 0.70, invisible(">"), invisible("<"))` 0.70) [Table S4]. This lack of difference can be highlighted when we visualize the results for butyrate kinase [Figure 2B]. These observations provide evidence that gene prevalence for enzymes involved in SCFA synthesis does not change due to colorectal tumors.

>>>>

**Decreased SCFA concentrations are not associated with adenoma or carcinoma tumors.** We used high-performance liquid chromatography (HPLC) to measure acetate, butyrate, and propionate concentrations of frozen fecal samples from `r length(metai$sample)` individuals at a cross-sectional point in time. There was no difference between individuals without colon tumors (n=`r length(filter(metai, dx == "normal") %>% pull(sample))`) and patients with either an adenoma (n=`r length(filter(metai, dx == "adenoma") %>% pull(sample))`) or carcinoma (n=`r length(filter(metai, dx == "cancer") %>% pull(sample))`) for any of the SCFAs measured after multiple comparison correction (P-value `r ifelse(min(pull(hplc_groups_scfa_kruskal, bh)) > 0.15, invisible(">"), invisible("<"))` 0.15) [Figure 1A - 1C]. We next measured the concentration of SCFAs in `r length(metaf$initial)` patients with an adenoma (n=`r length(filter(metaf, dx == "adenoma") %>% pull(initial))`) or carcinoma (n=`r length(filter(metaf, dx == "cancer") %>% pull(initial))`) in which we had pre- and post-treatment fecal samples. Although there was a general trend for increasing acetate, butyrate, and propionate concentrations following treatment for tumors, there was no significant difference between pre- and post-treatment for either patients with adenomas (P-value `r ifelse(min(filter(hplc_treat_scfa_kruskal, disease == "adenoma") %>% pull(bh)) > 0.20, invisible(">"), invisible("<"))` 0.20) or carcinomas (P-value `r ifelse(min(filter(hplc_treat_scfa_kruskal, disease == "carcinoma") %>% pull(bh)) > 0.80, invisible(">"), invisible("<"))` 0.80) [Figure 1D - 1F].

**Gene abundance for enzymes involved in SCFA synthesis are the same for individuals without tumors and patients with adenomas or carcinomas.** In order to provide further confirmation and support of our SCFA concentration results we investigated the genes encoding specific enzymes involved with SCFA synthesis [Table S1]. Using this list of specific genes [Table S1], we looked for differences in gene abundance between individuals without colon tumors and patients with adenomas or carcinomas. Although we intended to analyze all the genes in the list [Table S1], not all of the KEGG genes were identified during our analysis. We first analyzed imputed gene relative abundance, calculated from our OTU data that was generated from 16S rRNA gene sequencing. We found no difference in any of the imputed gene relative abundances for enzymes involved with acetate, butyrate, or propionate synthesis (P-value `r ifelse(min(pull(picrust_select, bh)) > 0.90, invisible(">"), invisible("<"))` 0.90) [Table S2]. Since butyrate is one of the most studied SCFAs in the context of CRC, we visualized the observed lack of difference between indviduals without tumors and patients with adenomas or carcinomas in gene abundance for enzymes involved with SCFA synthesis using the butyrate kinase gene [Figure 2A]. Additionally, using a paired Wilcoxon rank-sum test, there also was no difference in imputed gene relative abundance between pre- and post-treatment samples for any genes involved with SCFA synthesis (P-value `r ifelse(min(pull(picrust_select_treat, bh)) > 0.70, invisible(">"), invisible("<"))` 0.70) [Table S3]. Next, we took a subset of these `r length(metai$sample)` fecal samples (n=`r length(geof_samples$X1)`) and used metagenomic sequencing to confirm these results. Like the imputed gene results, metagenomic analysis found that there was no difference in any of the genes involved in SCFA synthesis between individuals without colon tumors (n=`r length(filter(geof_samples, X30 == "Healthy") %>% pull(X2))`) and patients with adenomas (n=`r length(filter(geof_samples, X30 == "Adenoma") %>% pull(X2))`) or carcinomas (n=`r length(filter(geof_samples, X30 == "Cancer") %>% pull(X2))`) (P-value `r ifelse(min(pull(opf_select, bh)) > 0.70, invisible(">"), invisible("<"))` 0.70) [Table S4]. This similarity between indviduals without tumors and patients with adenomas or carcinomas is highlighted again by visualizing the gene abundance for the butyrate kinase gene [Figure 2B]. These observations provide evidence that gene prevalence for enzymes involved in SCFA synthesis does not change due to colorectal tumors and provides further support for our original SCFA concentration observations.

### 2018-08-03 submission/manuscript.tex
include Ada suggestions results 1st 2 paragraphs

\textbf{Decreased SCFA concentrations are not associated with tumors.}
We used high-performance liquid chromatography (HPLC) to measure
acetate, butyrate, and propionate concentrations of frozen fecal samples
from 490 individuals at a cross-sectional point in time. There was no
difference between individuals without colon tumors (n=172) and patients
with either an adenoma (n=198) or carcinoma (n=120) for any of the SCFAs
measured after multiple comparison correction (P-value \textgreater{}
0.15) {[}Figure 1A - 1C{]}. We next measured the concentration of SCFAs
in 67 patients with an adenoma (n=41) or carcinoma (n=26) in which we
had pre- and post-treatment fecal samples. Although there was a general
trend for increasing acetate, butyrate, and propionate concentrations
after treatment for tumors, there was no significant difference pre- and
post-treatment for patients with adenomas (P-value \textgreater{} 0.20)
or carcinomas (P-value \textgreater{} 0.80) {[}Figure 1D - 1F{]}.

\textbf{Changes in genes for enzymes involved in SCFA synthesis are not
associated with tumors.} Using a list of specific genes that are
important for the synthesis of SCFAs {[}Table S1{]}, we looked for
differences in gene abundance between individuals without colon tumors
and patients with adenomas or carcinomas. First, using imputed gene
relative abundance based on 16S rRNA gene sequencing we found no
difference in any of the genes involved with acetate, butyrate, or
propionate synthesis (P-value \textgreater{} 0.90) {[}Table S2{]}. This
similarity between groups is highlighted by visualizing genes important
in butyrate synthesis {[}Figure 2A{]}. Using a paired Wilcoxon rank-sum
test, there also was no difference in imputed gene relative abundance
between pre- and post-treatment samples for any genes involved with SCFA
synthesis (P-value \textgreater{} 0.70) {[}Table S3{]}. Next, we took a
subset of these 490 fecal samples (n=85) and used metagenomic sequencing
to confirm these results. Like the imputed gene results, metagenomic
analysis found that there was no difference in any of the genes involved
in SCFA synthesis between individuals without colon tumors (n=29) and
patients with adenoma (n=28) or carcinoma (n=28) (P-value \textgreater{}
0.70) {[}Table S4{]}. This lack of difference can be highlighted when we
visualize the results for butyrate kinase {[}Figure 2B{]}. These
observations provide evidence that gene prevalence for enzymes involved
in SCFA synthesis does not change due to colorectal tumors.

>>>>

\textbf{Decreased SCFA concentrations are not associated with adenoma or
carcinoma tumors.} We used high-performance liquid chromatography (HPLC)
to measure acetate, butyrate, and propionate concentrations of frozen
fecal samples from 490 individuals at a cross-sectional point in time.
There was no difference between individuals without colon tumors (n=172)
and patients with either an adenoma (n=198) or carcinoma (n=120) for any
of the SCFAs measured after multiple comparison correction (P-value
\textgreater{} 0.15) {[}Figure 1A - 1C{]}. We next measured the
concentration of SCFAs in 67 patients with an adenoma (n=41) or
carcinoma (n=26) in which we had pre- and post-treatment fecal samples.
Although there was a general trend for increasing acetate, butyrate, and
propionate concentrations following treatment for tumors, there was no
significant difference between pre- and post-treatment for either
patients with adenomas (P-value \textgreater{} 0.20) or carcinomas
(P-value \textgreater{} 0.80) {[}Figure 1D - 1F{]}.

\textbf{Gene abundance for enzymes involved in SCFA synthesis are the
same for individuals without tumors and patients with adenomas or
carcinomas.} In order to provide further confirmation and support of our
SCFA concentration results we investigated the genes encoding specific
enzymes involved with SCFA synthesis {[}Table S1{]}. Using this list of
specific genes {[}Table S1{]}, we looked for differences in gene
abundance between individuals without colon tumors and patients with
adenomas or carcinomas. Although we intended to analyze all the genes in
the list {[}Table S1{]}, not all of the KEGG genes were identified
during our analysis. We first analyzed imputed gene relative abundance,
calculated from our OTU data that was generated from 16S rRNA gene
sequencing. We found no difference in any of the imputed gene relative
abundances for enzymes involved with acetate, butyrate, or propionate
synthesis (P-value \textgreater{} 0.90) {[}Table S2{]}. Since butyrate
is one of the most studied SCFAs in the context of CRC, we visualized
the observed lack of difference between indviduals without tumors and
patients with adenomas or carcinomas in gene abundance for enzymes
involved with SCFA synthesis using the butyrate kinase gene {[}Figure
2A{]}. Additionally, using a paired Wilcoxon rank-sum test, there also
was no difference in imputed gene relative abundance between pre- and
post-treatment samples for any genes involved with SCFA synthesis
(P-value \textgreater{} 0.70) {[}Table S3{]}. Next, we took a subset of
these 490 fecal samples (n=85) and used metagenomic sequencing to
confirm these results. Like the imputed gene results, metagenomic
analysis found that there was no difference in any of the genes involved
in SCFA synthesis between individuals without colon tumors (n=29) and
patients with adenomas (n=28) or carcinomas (n=28) (P-value
\textgreater{} 0.70) {[}Table S4{]}. This similarity between indviduals
without tumors and patients with adenomas or carcinomas is highlighted
again by visualizing the gene abundance for the butyrate kinase gene
{[}Figure 2B{]}. These observations provide evidence that gene
prevalence for enzymes involved in SCFA synthesis does not change due to
colorectal tumors and provides further support for our original SCFA
concentration observations.
